---
title: "FML assignment 2"
author: "Saride Lakshmi Priya"
date: "2023-10-01"
output:
pdf_document: default
html_document: default
word_document: default
---

## Data import 
#First load the required libraries
```{r}
library(class)

library(caret)
```


```{r}
library(e1071)
```

# Read the file data
```{r}
data<-read.csv("C:/Users/DELL/Downloads/UniversalBank.csv")
dim(data)

```
```{r}
head(data)
```

```{r}
tail(data)
```
```{r}
t(t(names(data))) 
```
# drop ID & ZIP
```{r}
new_df <-data[,-c(1,5)]
dim(new_df)
```
#Education needs to be converted to factor

```{r}
new_df$Education <- as.factor(new_df$Education)
```

# creating education to dummy variables
```{r}
dummy_var <- dummyVars(~.,data=new_df)
predict_data <- as.data.frame(predict(dummy_var,new_df))

```

#When we need to run the code again, we partition the data into training (60%) and validation (40%) sets and set the seed.

```{r}
set.seed(1)
train.data <- sample(row.names(predict_data), 0.6*dim(predict_data)[1])
valid.data <- setdiff(row.names(predict_data),train.data)
train_data <- predict_data[train.data,]
valid_data <- predict_data[valid.data,]
t(t(names(train_data)))

```
```{r}
summary(train_data)
```

```{r}
cat("The size of the training dataset is:",nrow(train_data))

```

```{r}
summary(valid_data)

```

```{r}
cat("The size of the validation dataset is:",nrow(valid_data))

```
Normalizing of the data

```{r}
train.norm <- train_data[,-10]
valid.norm <- valid_data[,-10]
norm <- preProcess(train_data[,-10],method=c("center","scale"))
train.norm <- predict(norm,train_data[,-10])
valid.norm <- predict(norm,valid_data[,-10])
```

Questions

Consider the following customer:

1. Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities
Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform
a k-NN classification with all predictors except ID and ZIP code using k = 1.
Remember to transform categorical predictors with more than two categories
into dummy variables first. Specify the success class as 1 (loan acceptance), and
use the default cutoff value of 0.5. How would this customer be classified?

Creating new customer data

```{r}
new.cust <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1
)
# Normalize the new customer dataset
cust.norm <- predict(norm, new.cust)
```

let us predict using KNN

```{r}
prediction <- class::knn(train = train.norm,
test = cust.norm,
cl = train_data$Personal.Loan, k = 1)
prediction
```
2.What is a choice of k that balances between overfitting and ignoring the predictor information?

```{r}
# Calculate the accuracy for each value of k
# Set the range of k values to consider
accuracy <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
knn <- class::knn(train = train.norm,
test = valid.norm,
cl = train_data$Personal.Loan, k = i)
accuracy[i, 2] <- confusionMatrix(knn,
as.factor(valid_data$Personal.Loan),positive = "1")$overall[1]
}
which(accuracy[,2] == max(accuracy[,2]))

```
```{r}
accuracy
```

```{r}
plot(accuracy$k,accuracy$overallaccuracy)
```
3. Show the confusion matrix for the validation data that results from using the best k.


```{r}
prediction1 <- class::knn(train = train.norm,
test = valid.norm,
cl = train_data$Personal.Loan, k=3)
confusionMatrix(prediction1,as.factor(valid_data$Personal.Loan))

```
4. Consider the following customer: Age = 40, Experience = 10, Income =
84,Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3
= 0,Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and
CreditCard = 1. Classify the customer using the best k.

now building the second new customer dataset

```{r}
customer2.df <- data.frame(
Age = 40,
Experience = 10,
Income = 84,
Family = 2,
CCAvg = 2,
Education.1 = 0,
Education.2 = 1,
Education.3 = 0,
Mortgage = 0,
Securities.Account = 0,
CD.Account = 0,
Online = 1,
CreditCard = 1)
#Normalizing the 2nd customer dataset
cust_norm2 <- predict(norm , customer2.df)

```

Question-5: Repeating the process by partitioning the data into three parts -
50%, 30%, 20%,Apply the k-NN method with the k chosen above. Compare the
confusion matrix of the test set with that of the training and validation sets.
Comment on the differences and their reason.

```{r}
set.seed(500)
Train_Index <- sample(row.names(predict_data), .5*dim(predict_data)[1])#create train index
```

```{r}
#create validation index
Val_Index <- sample(setdiff(row.names(predict_data),Train_Index),.3*dim(predict_data)[1])
Test_Index =setdiff(row.names(predict_data),union(Train_Index,Val_Index))#create test index
train.df <- predict_data[Train_Index,]
cat("The size of the new training dataset is:", nrow(train.df))
## The size of the new training dataset is: 2500
valid.df <- predict_data[Val_Index, ]
cat("The size of the new validation dataset is:", nrow(valid.df))
## The size of the new validation dataset is: 1500
test.df <- predict_data[Test_Index, ]
cat("The size of the new test dataset is:", nrow(test.df))
```
 
 Normalizing the data
 
```{r}
norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.df.norm <- predict(norm.values, train.df[, -10])
valid.df.norm <- predict(norm.values, valid.df[, -10])
test.df.norm <- predict(norm.values, test.df[,-10])

```

Performing kNN and creating confusion matrix on training, testing, validation
data

```{r}
prediction2 <- class::knn(train = train.df.norm,
test = test.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(prediction2,as.factor(test.df$Personal.Loan))
```

```{r}
prediction3 <- class::knn(train = train.df.norm,
test = valid.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(prediction3,as.factor(valid.df$Personal.Loan))
```
```{r}
prediction4 <- class::knn(train = train.df.norm,
test = train.df.norm,
cl = train.df$Personal.Loan, k=3)
confusionMatrix(prediction4,as.factor(train.df$Personal.Loan))

```

